---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Bis557

<!-- badges: start -->
![Travis build status](https://travis-ci.com/importbq/bis557.svg?token=XyjFaGxnEDaFXgSHunRQ&branch=master)
[![codecov](https://codecov.io/gh/importbq/bis557/branch/master/graph/badge.svg)](https://codecov.io/gh/importbq/bis557)
<!-- badges: end -->


This Package contains my homeworks for Bis 557

## Installation

The development version of the packages

``` r
install.packages("bis557")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("importbq/bis557")
```
## Example

We have an example here to show the usage of linear_model:

```{r example}
library(bis557)
data(iris)
fit_linear_model <- linear_model(Sepal.Length ~ ., iris)
print(fit_linear_model)
```
We see that gradient descent achieved very good results for OLS.

# Please read Vignette for details

# HW2 Vignette has the solution to HW2
